---
title: "Analysis Walkthrough"
author: "Hayley Lenamon"
format:
  html:
    self-contained: true
    code-fold: true
    code-tools: true
---

To begin, I loaded the necessary libraries. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#Load libraries 
library(tidyverse) 
library(janitor) 
library(readxl) 
library(scales)
library(lubridate)
library(tidyverse)
library(tigris)
library(sf)
library(tmap)
library(tmaptools)
library(htmltools)
library(janitor)
library(rmapshaper)
library(here)
options(tigris_class = "sf")
```

The data used for this assignment includes Georgia election results by county for U.S. Senate and Governor in 2022, President in 2020, and then some other socioeconomic factors.

The first step is to load in the data, which is stored in the `ga_votes.xlsx` Excel file that's included in this repository. There are two tabs - one contains the senate results, and one contains the governor results. Below is the information I used to import the GA election data. 

```{r}
senate_results <- read_excel(here("ga_votes.xlsx"), sheet = "senate")

governor_results <- read_excel(here("ga_votes.xlsx"), sheet = "governor")
```

I created a column/bar chart for the Senate race that shows how many counties the Republican candidate did better vs. worse than Trump.

```{r}
ggplot(data=senate_results, aes(x=factor(is_r_above_trump))) + geom_bar()

```

Then I did the same as Question 1, just for the race for Governor.

```{r}
ggplot(data=governor_results, aes(x=factor(is_r_above_trump))) + geom_bar()

```

For the Senate race again, I created a chart that shows the top 10 biggest drops between the Republican candidate and Trump (i.e. the 10 counties with the most substantial negative numbers in `pct_r_above_trump`). I sliced off a certain number of rows using the `head()` function. 

```{r}
no3 <- arrange(senate_results, pct_r_above_trump)
top10senate <- no3 %>% head(10)

ggplot(data=top10senate, aes(x=factor(pct_r_above_trump))) + geom_bar() 
```

Then I created a chart that shows the top 10 counties highest margins of victory for the Democratic candidate for Senate. To do that I compared the Democratic candidate's percentage (pct_d\_2022) with the Republican's (pct_r\_2022) to calculate the margin of victory for each county.

```{r}
senate_margin <- senate_results %>% mutate(dem_margin = pct_d_2022 - pct_r_2022)
no4 <- arrange(senate_margin, dem_margin)
top10_senate_margin <- no4 %>% head(10)

ggplot(data=top10_senate_margin, aes(x=factor(dem_margin))) + geom_bar()

```

Then I did the same as Q4, only this time for the race for Governor.

```{r}
gov_margin <- governor_results %>% mutate(dem_margin = pct_d_2022 - pct_r_2022)
no5 <- arrange(gov_margin, dem_margin)
top10_gov_margin <- no5 %>% head(10)

ggplot(data=top10_gov_margin, aes(x=factor(dem_margin))) + geom_bar()

```
