---
title: "Analysis Walkthrough"
author: "Hayley Lenamon"
format:
  html:
    self-contained: true
    code-fold: true
    code-tools: true
---
To begin, I loaded the necessary libraries. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(nycflights13) # <- make sure to install this package 
library(tigris)
library(tidycensus)
library(tmap)
library(tmaptools)
library(janitor)
library(writexl)
library(here)
options(tigris_class = "sf")
```

# Functions

I first created a function that returns the sum of 2 numbers.

```{r}
sum_function <- function(number_one, number_two){sum <- number_one + number_two
return(sum)}
sum_function(15,9384)
```

Next, I created a function that returns the mean of a list of numbers fed to it. 

```{r}
list_mean <- c(12,53,62,28,16)
resultmean <- mean(list_mean)
print(resultmean)
```

Now, I expanded on the function above and instead of just returning the mean number alone, it returns the sentence: "The mean is ___". (Where the blank is the mean.)

```{r}
list_mean <- c(12,53,62,28,16)

result_mean_print <- function(list_mean){
  result <- paste0("The mean is ", mean(list_mean), ".") 
  return(result)}

result_mean_print(list_mean)
```

For the next two questions, I use the `flights` dataset from the package `nycflights13`. It contains records of individual airline flights departing at New York's three metro airports, JFK, LaGuardia (LGA) and Newark (EWR). Run the following chunk to load the data. 

```{r}
#load the data to use for the rest of the assignment questions
flights <- nycflights13::flights
head(flights)
```
4) Using the `flights` dataset, I created a function that filters the data frame to only include records from a single originating airport (column name `origin`). 

```{r}
flight_df <- function(origin_name){records <- flights %>%
  filter(origin == origin_name) 
return(records)}
flight_df("JFK")
```

Similarly, using the `flights` dataset, I created a function that filters the data frame to only include records from a single originating airport (column name `origin`), and then aggregates (i.e. counts) the results by airline (`carrier`) to show how many flights each airline has from that airport.

```{r}
airport_df <- function(origin_name){
  records <- flights %>%
    filter(origin == origin_name) %>%
    group_by(carrier) %>%
    count(carrier)
  return(records)}
airport_df("JFK")
```
